---
title: "RCT Step 5"
format: html
editor: visual
---

## 1. Comparisons

::: {.callout-note appearance="simple" icon="false"}
## Instructions

Compare the total assets and equity ratios (`equity / total assets`)

-   for all firms from your postal code
-   with those of the entire Berlin population of firms
:::

Load and check data

```{r}
#| echo: true
#| eval: true
#| results: hide
# Load data
orbis_data <- readRDS("~/Documents/rct25/data/generated/orbis_panel_berlin.rds")

# Check if data really only includes Berlin zip codes
table(orbis_data$city_native)

# Check how many NAs
mean(is.na(orbis_data$postcode)) * 100
mean(is.na(orbis_data$toas)) * 100
mean(is.na(orbis_data$shfd)) * 100

```

Prepare comparison

```{r}
# Focus on 2021 for comparability
orbis_2021 <- subset(orbis_data, year == 2021)

# Create two subsets: 
# "plänterwald" (zip code 12437) and "berlin" (all zip codes except 12437)
zip_plänterwald_firms <- subset(orbis_2021, postcode == 12437)
zip_berlin_firms <- subset(orbis_2021, postcode != 12437)

# Create new column for equity ratio
# equity ratio = equity / total assets
zip_plänterwald_firms$equity_ratio <- zip_plänterwald_firms$shfd / zip_plänterwald_firms$toas
zip_berlin_firms$equity_ratio <- zip_berlin_firms$shfd / zip_berlin_firms$toas

# Calculate means
mean_assets_plänterwald <- mean(zip_plänterwald_firms$toas, na.rm = TRUE)
mean_assets_berlin <- mean(zip_berlin_firms$toas, na.rm = TRUE)
mean_eq_plänterwald <- mean(zip_plänterwald_firms$equity_ratio, na.rm = TRUE)
mean_eq_berlin <- mean(zip_berlin_firms$equity_ratio, na.rm = TRUE)

# Calculate sds
sd_assets_plänterwald <- sd(zip_plänterwald_firms$toas, na.rm = TRUE)
sd_assets_berlin <- sd(zip_berlin_firms$toas, na.rm = TRUE)
sd_eq_plänterwald <- sd(zip_plänterwald_firms$equity_ratio, na.rm = TRUE)
sd_eq_berlin <- sd(zip_berlin_firms$equity_ratio, na.rm = TRUE)
```

Show comparison in table

```{r, results='asis'}
#| echo: true
#| message: false
#| warning: false

library(dplyr)
library(kableExtra)
library(knitr)
library(scales)

# Create dataframe (make sure rounding and format is correct)
table_data <- data.frame(
  Variable = c("Total Assets", "Equity Ratio"),
  `Plänterwald Mean` = c(comma(round(mean_assets_plänterwald, 2)), round(mean_eq_plänterwald, 2)),
  `Plänterwald SD` = c(comma(round(sd_assets_plänterwald, 2)), round(sd_eq_plänterwald, 2)),
  `Berlin Mean` = c(comma(round(mean_assets_berlin, 2)), round(mean_eq_berlin, 2)),
  `Berlin SD` = c(comma(round(sd_assets_berlin, 2)), round(sd_eq_berlin, 2))
)

# Create layout for the table
table_html <- table_data %>%
  kable(
    format = "html",
    escape = FALSE,
    digits = 2,
    caption = "Table 1: Comparison of Total Assets and Equity Ratio (2021)",
    col.names = c("Variable", "Mean", "SD", "Mean", "SD"),
    align = "lcccc"
  ) %>%
  add_header_above(c(" " = 1, "Plänterwald" = 2, "Berlin (Rest)" = 2)) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  as.character()

note_html <- "<div style='margin-top: 10px; font-size: 0.9em; color: #555;'>
<strong>Note:</strong> This table shows summary statistics for firms located in Plänterwald (n = 48) compared to the rest of Berlin (n = 27,688) in fiscal year 2021, based on Orbis data. Total assets are reported in euros (€) with comma formatting. The equity ratio is calculated as shareholders’ funds divided by total assets, expressed as a percentage. Reported are group means and standard deviations (SD) excluding missing values (NA). All figures are based on the most recent available financial data for 2021.
</div>"

cat(table_html, note_html, sep = "\n")
```

## 2. Statistically significant differences

::: {.callout-note appearance="simple" icon="false"}
## Instructions

Test for statistically significant differences.
:::

T-test for total assets and equity ratio

```{r}
# Berechne T-Tests
t_assets <- t.test(zip_plänterwald_firms$toas, zip_berlin_firms$toas, alternative = "two.sided", var.equal = FALSE)
t_eq <- t.test(zip_plänterwald_firms$equity_ratio, zip_berlin_firms$equity_ratio, alternative = "two.sided", var.equal = FALSE)

# prepare all values with correct rounding and formatting
table_data <- data.frame(
  Variable = c("Total Assets", "Equity Ratio"),
  `Plänterwald Mean` = c(comma(round(mean_assets_plänterwald, 2)), round(mean_eq_plänterwald, 2)),
  `Plänterwald SD` = c(comma(round(sd_assets_plänterwald, 2)), round(sd_eq_plänterwald, 2)),
  `Berlin Mean` = c(comma(round(mean_assets_berlin, 2)), round(mean_eq_berlin, 2)),
  `Berlin SD` = c(comma(round(sd_assets_berlin, 2)), round(sd_eq_berlin, 2)),
  `t stat` = round(c(t_assets$statistic, t_eq$statistic), 3),
  `df` = round(c(t_assets$parameter, t_eq$parameter), 0),
  `95% CI Lower` = c(comma(round(t_assets$conf.int[1], 2)), round(t_eq$conf.int[1], 2)),
  `95% CI Upper` = c(comma(round(t_assets$conf.int[2], 2)), round(t_eq$conf.int[2], 2)),
  `p-value` = c(
    ifelse(t_assets$p.value < 0.001, "<0.001", round(t_assets$p.value, 3)),
    ifelse(t_eq$p.value < 0.001, "<0.001", round(t_eq$p.value, 3))
  )
)

```

-\> will be added to first table in step 4

## 3. Research Log

::: {.callout-note appearance="simple" icon="false"}
## Instructions

Document and explain your research design decisions in your research log.
:::

-\> This can be found in research log file.

## 4. Table

::: {.callout-note appearance="simple" icon="false"}
## Instructions

Prepare an informative, self-explanatory table (including notes) that clearly communicates your findings.
:::

```{r results='asis'}
library(kableExtra)
library(dplyr)
library(htmltools)

table_html <- table_data %>%
  kable(
    format = "html",
    escape = FALSE,
    digits = 2,
    caption = "Table 1: Comparison of Total Assets and Equity Ratio (2021)",
    col.names = c("Variable", "Mean", "SD", "Mean", "SD", "t stat", "df", "95% CI Lower", "95% CI Upper", "p-value"),
    align = "lcccccclcc"
  ) %>%
  add_header_above(c(" " = 1, "Plänterwald" = 2, "Berlin" = 2, "T-Test" = 5)) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  as.character()

note_html <- "<div style='margin-top: 10px; font-size: 0.9em; color: #555;'>
<strong>Note:</strong> Notes: This table presents descriptive statistics for firms located in Plänterwald (n = 48) compared to the remaining Berlin-based firms (n = 27,688) for the fiscal year 2021, based on Orbis data. Total assets are reported in euros (€) and formatted using comma separators. The equity ratio is defined as shareholders’ funds divided by total assets, expressed as a percentage. Reported are means and standard deviations (SD) for both groups. Welch’s two-sample t-test was employed to compare group means, accounting for unequal variances. Reported statistics include t-values, degrees of freedom (df), p-values, and 95% confidence intervals (CI). The large SD in Berlin firms’ total assets and equity ratios likely reflects high variance, potentially due to outliers or distressed companies. All values are based on financial data from the year 2021.
</div>"

cat(table_html, note_html, sep = "\n")
```

PDF export of Table 2

```{r}
#| echo: true
#| eval: true

# Create second HTML file that only shows final table
full_html <- paste0(
  "<!DOCTYPE html>",
  "<html><head><meta charset='utf-8'>",
  "<title>Table 1</title>",
  "<link rel='stylesheet' href='https://cdnjs.cloudflare.com/ajax/libs/twitter-bootstrap/4.5.2/css/bootstrap.min.css'>",
  "</head><body style='margin: 20px;'>",
  table_html,
  note_html,
  "</body></html>"
)

writeLines(full_html, "Table_Step5V2.html")

# Automatically export HTML as PDF
library(pagedown)
pagedown::chrome_print("Table_Step5.html", output = "Table_Step5.pdf")

```
